---
title: "Filter or split cevodata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Filter or split cevodata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Transformations

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cevodata)
```

3 cevodata transforming functions have been implemented:

- `filter()`
- `split_by()`
- `merge()`

## filter()

`filter()` can be used to subset the cevodata using any metadata columns in a tidyverse manner

```{r}
test_data |> 
  filter(sex == "female")
```

## split_by()

`split_by()` to split cevodata by any metadata column

```{r}
splits <- test_data |> 
  split_by(sex)

splits
```

```{r}
splits$male
```


## merge()

`merge()` to merge 2 objects into one

```{r}
merge(splits$male, splits$female)
```


Transforming functions are helpful in investigating particular samples, for example, when one wants to see the VAF spectrum of one sample only quickly:

```{r}
test_data |> 
  filter(sample_id == "Sample 2") |> 
  plot_SFS()
```

